---
title: "Calculating importance values"
output: html_notebook
---
## Introduction

This markdown document will walk through the calculation of importance values from first calculating frequency, density and dominance values for species in a forest plot.

The test data set in this project comes from the 2017 Forest Ecology class.

## Step 1. Set up the workspace
```{r}
rm(list = ls())
library(dplyr)
```

## Step 2. Pull in and review data.  Fix any errors

```{r}
trees<-read.csv("trees.csv", sep = ",", stringsAsFactors = F)
```
Now get rid of some columns we don't need

```{r}
library(dplyr)
delete_cols<-c("PlotNum", "PlotID", "Quadrant", "Xcoord", "Ycoord", "DistanceM", "HtPercent", "TreeHeightM")

trees<-trees[,!(names(trees)%in% delete_cols)]
```
Now let's make some factors
```{r}
trees$Forest<-as.factor(trees$Forest)
trees$Species<-as.factor(trees$Species)
```
Now check levels of factors
```{r}
levels(trees$Forest)
levels(trees$Species)
```
Need to fix some species issues:
1) `Sugar maple?`, `Sugar maple` are the same thing.
2) `not recorded`, `Unknown1` and `unknown` should be pooled

```{r}
acerSaccharum<-which(trees$Species == "Sugar maple?" | trees$Species == "Sugar maple")
unknown<-which(trees$Species == "not recorded" | trees$Species == "Unknown1" | trees$Species == "unknown")
```
Now fix
```{r}
trees$Species[acerSaccharum]<-"Sugar maple"
trees$Species[unknown]<-"unknown"

```
Now regain factor levels
```{r}
trees$Species<-factor(trees$Species)
levels(trees$Species)
```
And now get rid of the unknowns
```{r}
trees<-filter(trees, Species != "unknown")
```
Now omit dead trees
```{r}
dead_ones<-which(trees$Comments == "DEAD" | trees$Comments == "1/2 DEAD")

trees<-trees[-dead_ones,]
```


Now save data
```{r}
write.csv(trees, "cleanTrees.csv")
```


